Peer Assignment 1
======================================================
####Author:Carlos José Rivas
####Course: Reproducible Research-Coursera


Setting the Working Directory
```{r SettingWd, echo=TRUE}
setwd("C:/Users/CarlosJosé/Desktop/Data Mining/Jonhs Hopkins/Ciencia de Datos/Reproducible Research/Peer Asigment")  
```
**Reading the activity.csv**
```{r ReadCSV, echo=TRUE}
myTbl=read.csv("activity.csv", colClasses = c("numeric","Date","numeric"))
##Creating sub set with NA removed
dataNAOmitted= subset(myTbl, is.na(myTbl$steps)==F)
```
what is mean total number of steps taken per day?
1-Make a histogram of the total number of steps taken each day
```{r}
library(plyr)
TotalPerDay= ddply(dataNAOmitted, .(date), summarise, steps=sum(steps))
hist(TotalPerDay$steps,col="blue")
```
2.Calculate and report the mean and median total number of steps taken per day

```{r echo=TRUE}
##Mean
mean(TotalPerDay$steps)
##mediana
median(TotalPerDay$steps)
```
What is the average daily activity patter?
```{r echo=TRUE}
avgPerInterval= ddply(dataNAOmitted, .(interval),summarise, steps=mean(steps))
plot(avgPerInterval$interval, avgPerInterval$steps, type="l")
```
Which 5 minute interval, on average across all the days in the dataset contains the maximum number of steps

```{r echo=TRUE}
avgPerInterval[which.max(avgPerInterval$steps),]
```

Imputting missing values
```{r echo=TRUE}
##Cantidad de NA
sum(is.na(myTbl$steps))
##i will try to fill NA with average
##Creating new data set
imput= myTbl
for(i in 1:nrow(imput)){
        if (is.na(imput$steps[i])) {
           imput$steps[i]= avgPerInterval$steps[which(imput$interval[i]==avgPerInterval$interval)]
        }
}

imput= arrange(imput,interval)
TotPerdayImputed= ddply(imput, .(date),summarise, steps=sum(steps))
hist(TotPerdayImputed$steps,col="blue")
##media
mean(TotPerdayImputed$steps)
##mediana
median(TotPerdayImputed$steps)
##diferencias
abs(mean(TotalPerDay$steps)-mean(TotPerdayImputed$steps))
abs(median(TotalPerDay$steps)-median(TotPerdayImputed$steps))
##steps difference
GranDiff= sum(imput$steps)- sum(dataNAOmitted$steps)
GranDiff
```

Create a new factor variable in the data set with two levels weekday and weekend

```{r}
Sys.setlocale("LC_TIME", "English")
imput$weekdays=weekdays(as.Date(imput$date))
imput$weekdays= ifelse(imput$weekdays %in% c("Saturday", "Sunday"), "weekend", 
    "weekday")

myAvg= ddply(imput, .(interval,weekdays), summarise, steps=mean(steps))

library(lattice)
xyplot(steps~interval|weekdays, data=myAvg, layout=c(1,2),type="l")
```
























